Supplementary Material for ``Another Look at Regularized Discriminant Analysis for High-Dimensional Classification''

### 2013-04-21 reports/supplementary-material.Rnw
Moved figures of HDRDA optimal estimates to Supplementary Material document.

\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}

@@

opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE,
               warning = FALSE, cache = TRUE, echo = FALSE, autodep = TRUE,
               fig.width = 10, fig.height = 10, out.width = "\\linewidth")
@

@@

\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{thm}{Theorem}

\bibpunct{(}{)}{;}{a}{,}{,}

@@

\begin{abstract}

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, the analysis of microarray data often presents difficult
challenges because the number of gene expression levels is typically much
greater than the sample size. Consequently, several sparse and regularized
classifiers have been proposed, often with strict covariance assumptions or
eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
propose a novel modification of the regularized discriminant analysis classifier
popularized by \cite{Friedman:1989tm}. Our proposed classifier provides a
straightforward interpretation and employs a new dimension-reduction technique
to enable a computationally efficient model-selection procedure, thus overcoming
the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
high-dimensional data. Moreover, we demonstrate with several high-dimensional
microarray data sets that our proposed classifier is competitive and often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

\end{abstract}

\section{Introduction}

Microarrays have enabled researchers to measure gene-expression levels of
thousands of genes in a single experiment to identify a subset of genes and
their expression patterns that either characterize a particular cell state or
facilitate an accurate prediction of future cell state.  We focus on the
classification of tissues and other biological samples into one of $K$ disjoint,
predefined cancer types, where our goal is to ascertain the class membership of
a new tissue sample based on gene-expression microarray data. However,
microarray data typically consist of a large number of gene expressions $p$
relative to the total sample size $N$. Moreover, standard benchmark microarray
data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
than $N = 100$ observations, in which case standard supervised-learning methods
exhibit poor classification performance because of the \emph{curse of
  dimensionality} \citep{Bellman:1961tn}. Furthermore, some well-known
classifiers, such as linear and quadratic discriminant analysis, are
incalculable when $p > N$ because the sample covariance matrices employed are
singular. As a result, many researchers, including \cite{Mai:2012bf},
\cite{Fan:2012iq}, \cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and
\cite{Huang:2010ju}, have introduced restrictions on the population covariance
matrices in linear and quadratic discriminant analysis to reduce number of
covariance parameters that must be estimated to ensure that the resulting
classifiers are calculable. For instance, \cite{Dudoit:2002ev} have modified the
linear and quadratic discriminant analysis classifiers by assuming that the
population covariance matrices are diagonal, resulting in classifiers that are
special cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.

Alternatively, several researchers, including \cite*{Zhang:2010va},
\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
regularization methods to improve the estimation of the population covariance
matrices before applying either linear or quadratic discriminant
analysis. Covariance-matrix regularization methods often are based on the
well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. For excellent overviews of the regularization problem in
discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we propose an improvement to the popular regularized discriminant
analysis classifier of \cite{Friedman:1989tm}, who has proposed pooling of the
covariance-matrix estimators from linear and quadratic discriminant analysis and
has additionally incorporated a covariance-matrix shrinkage term to ensure that
the classifier is well-posed. \citeauthor{Friedman:1989tm}'s classifier can
attain improved classification accuracy, compared to linear and quadratic
discriminant analysis when $p < N$. However, in its original form,
\citeauthor{Friedman:1989tm}'s classifier is impractical when $p \gg N$ because
the decision rule requires the calculation of the inverses and determinants of
large $p \times p$ covariance-matrix estimators for each class. Furthermore,
\citeauthor{Friedman:1989tm}'s classifier has two tuning parameters that are
selected via cross-validation, which adds a substantial computational burden to
the model-selection process. \citeauthor{Friedman:1989tm}'s classifier is
ill-advised on modern data sets because the model selection requires $O(Kp^3)$
calculations for each tuning-parameter pair.

In this paper we modify \citeauthor{Friedman:1989tm}'s classifier by first
constructing the pooling matrix as a convex combination of the covariance-matrix
estimators employed in linear and quadratic discriminant analysis and
demonstrate that this simplification improves the interpretation of the pooling
operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
assumption of equal population covariance matrices within linear discriminant
analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
proposed classifier can relax the linearity assumption by instead incorporating
unequal covariance matrices while retaining linear and quadratic discriminant
analysis as special cases. We then establish our classifier by applying
covariance-matrix regularization with a parameterization that includes a large
family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
\cite{Xu:2009fl}.

We show that our proposed classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
decision rule of our classifier and show that the sum related to the null space
of the pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of \citeauthor{Friedman:1989tm}'s classifier and provides a substantial increase
in computational efficiency and model selection. In fact, we show that the
inverse and determinants can be reduced to calculations of $n_k \times n_k$
matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
K)$. This reduction in calculation is clearly advantageous because $n_k$ is
often quite small in microarray data sets (e.g., 15--20 observations).

We also demonstrate that our proposed classifier is competitive with and often
superior to several recent classifiers designed for small-sample,
high-dimensional data. In particular, we compare our classifier with diagonal
linear discriminant analysis from \cite{Dudoit:2002ev} and the recently proposed
classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
of the competing classifiers using four small-sample, high-dimensional
benchmark microarray data sets.

\section{Regularized Discriminant Analysis}
\label{sec:rda}

\subsection{Preliminaries}
In discriminant analysis we wish to assign correctly an unlabeled
$p$-dimensional observation vector $x$ to one of $K$ unique, known classes
(or populations) by constructing a classifier from $N$ training
observations. Let $x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
  1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
$y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$
denotes the matrix space of all $a \times b$ matrices over the real field
$\mathbb{R}$. We assume that $(x_i, y_i)$ is a realization from a mixture
distribution $p(x) = \sum_{k=1}^K p(x | \omega_k) p(\omega_k)$, where
$p(x | \omega_k)$ is the probability density function of the $k$th class and
$p(\omega_k)$ is the prior probability of class membership of the $k$th
class. We further assume that each of $p(\omega_k)$ is equal $(k = 1, \ldots,
K)$.

Quadratic discriminant analysis is the optimal Bayesian decision rule with
respect to a $0-1$ loss function when $p(x | \omega_k)$ is the probability
density function of the multivariate normal distribution with known mean vector
$\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\Sigma_k
\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
matrices. Because $\mu_k$ and $\Sigma_k$ are typically unknown, we assign
an unlabeled observation $x$ to class $\omega_k$ with the sample quadratic
discriminant analysis classifier
\begin{align}
	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\Sigma}_k$ are the maximum likelihood
estimators for $\mu_k$ and $\Sigma_k$, respectively, and $A^{T}$ denotes the
matrix transpose of $A \in \mathbb{R}_{m \times p}$. If we assume
further that $\Sigma_k = \Sigma$ $(k = 1, \ldots, K)$, then the pooled
sample covariance matrix $\widehat{\Sigma}$ is substituted for $\widehat{
  \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\Sigma}_k \label{eq:pooled-cov}
\end{align}
is the maximum likelihood estimator for $\Sigma$. In this case \eqref{eq:qda}
reduces to the sample linear discriminant analysis classifier, and the
log-determinant is omitted because it is constant across all classes.

Throughout the document we also require the following additional notation. Let
$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be
the $m \times p$ matrix of zeros, such that $0_m$ is understood to denote
$0_{m \times m}$. Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne
0$ and $0$ otherwise. Let $A \oplus B$ denote the direct sum of $A \in
\mathbb{R}_{r \times r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
  1]{Lutkepohl:1996uz}. Let $A^+$ and $\mathcal{N}(A) = \{z \in \mathbb{R}_{p
  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and null space
of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{y
\in \mathbb{R}_{p \times 1} : y^{T}v = 0\ \forall v \in V \}$ denote
the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.

\subsection{Covariance Matrix Regularization}

The smallest eigenvalues and the directions associated with their eigenvectors
highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{
  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
then rank$(\widehat{\Sigma}_k) \le n_k$, which implies that at least $p -
n_k$ eigenvalues of $\widehat{\Sigma}_k$ are 0. Furthermore, although more
feature information is available to discriminate among the $K$ classes, if $p >
n_k$, \eqref{eq:qda} is incalculable because $\widehat{\Sigma}_k^{-1}$ does
not exist.

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\Sigma}_k(\gamma) =  \widehat{\Sigma}_k + \gamma I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\Sigma}_k$ towards $I_p$, thereby increasing the
eigenvalues of $\widehat{\Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

\subsection{Friedman's Regularized Discriminant Analysis Classifier}

\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
incorporating a weighted average of the sample covariance matrix $\widehat{
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\Sigma}$ to
estimate $\Sigma_k$ with
\begin{align}
  \widehat{\Sigma}_k(\lambda) = n_k^{-1}(\lambda) S_k(\lambda) \quad (k =
  1, \ldots, K),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
S_k = n_k \widehat{\Sigma}_k$, $S = \sum_{k=1}^K S_k$, and $
S_k(\lambda) = (1 - \lambda) S_k + \lambda S$. We can interpret
\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
$\widehat{\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $
\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
$\widehat{\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.

To further improve the estimation of $\Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
	\widehat{\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\Sigma}_k(\lambda)\}}{p} I_p,\label{eq:sig-rda}
\end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
$\lambda$ controls the amount borrowed from $\widehat{\Sigma}$ to estimate
$\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
amount of eigenvalue shrinkage applied. Finally, by substituting
\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
analysis classifier
\begin{align}
	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}
If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
permits a numerically stable matrix inversion.

\section{High-Dimensional Regularized Discriminant Analysis}

\subsection{Improvements to \citeauthor{Friedman:1989tm}'s Classifier}

Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
first simplifying the formulation of $\widehat{\Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define the convex combination
\begin{align}
  \widehat{\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\Sigma}_k + \lambda
  \widehat{\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
clear interpretation of $\widehat{\Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\Sigma}_k(\lambda) &= \widehat{\Sigma}_k + \lambda (\widehat{\Sigma} - \widehat{\Sigma}_k) \nonumber\\
%	&= \widehat{\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\Sigma}_{k'}  - \widehat{\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimate $\Sigma_k$. Hence, for $0 < \lambda
\le 1$, we estimate $\Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\Sigma}_k(\lambda)$ from
rank$(\widehat{\Sigma}_k)$ to rank$(\widehat{\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\Sigma_k$,
corresponding to $\widehat{\Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\Sigma}$. For brevity, when $\lambda = 1$, we define $X =
[\sqrt{c_{1k}(1)} x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} x_N^{T}]^{T}$ such that
$\widehat{\Sigma} = X^{T} X$. Similarly, for $\lambda = 0$, we define
$X_k = [\sqrt{c_{1k}(0)} x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} x_N^{T}]^{T}$ such
that $\widehat{\Sigma}_k = X_k^{T} X_k$. In Figure
\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
approaches 1, the contours become more similar, resulting in identical contours
for $\lambda = 1$.

\[ \left[\text{Figure \ref{fig:hdrda-contours} goes about here }\right]\]

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\Sigma}_k = \alpha_k \widehat{\Sigma}_k(\lambda) + \gamma I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
we define the high-dimensional regularized discriminant analysis classifier
\begin{align}
	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\Sigma}_k$ is nonsingular such that
$\tilde{\Sigma}_k^{-1}$ can be substituted for $\tilde{\Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\Sigma}_k$.

\subsection{Enhanced Computational Efficiency}

Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
itself to a more efficient calculation and a rapid model selection of the
parameters $\lambda$ and $\gamma$\footnote{TODO: Briefly mention the updating
  trick employed in the R package}. We decompose \eqref{eq:hdrda} into a sum of
two components, where the first summand consists of matrix operations applied to
low-dimensional matrices and the second summand corresponds to the null space of
$\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\Sigma}$ yield constant
quadratic forms across all classes and can be omitted. For $p \gg N$, the
constant component involves determinants and inverses of high-dimensional
matrices, and by ignoring these calculations, a substantial reduction in
computational costs is achieved relative to Friedman's classifier in
\eqref{eq:rda}\footnote{TODO: Add a brief comparison of the computations of
  \emph{RDA} and \emph{HDRDA} in terms of Big Oh notation}. Furthermore, a
byproduct is that adjustments to the associated eigenvalues have no effect on
\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
calculate efficiently the eigenvalue decomposition of $\widehat{\Sigma}$,
further reducing the computational costs of our proposed classifier.

First, we require the following relationship regarding the null spaces of
$\widehat{\Sigma}_k(\lambda)$, $\widehat{\Sigma}$, and $\widehat{
  \Sigma}_k$.

\begin{lemma}\label{lemma:null-spaces}
Let $\widehat{\Sigma}_k$ and $\widehat{\Sigma}$ be the maximum
likelihood estimators of $\Sigma_k$ and $\Sigma$, respectively. Let
$\widehat{\Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
  \Sigma}) \subset \mathcal{N}(\widehat{\Sigma}_k)$ $(k = 1, \ldots, K)$.
\end{lemma}
\begin{proof}
Let $z \in \mathcal{N}\{\widehat{\Sigma}_k(\lambda)\}$ for some $k = 1,
\ldots, K$. Hence, $0 = z^{T} \widehat{\Sigma}_k(\lambda) z = (1 -
\lambda) z^{T} \widehat{\Sigma}_k z + \lambda z^{T} \widehat{
  \Sigma} z$. Because $\widehat{\Sigma}_k, \widehat{\Sigma}\in
\mathbb{R}_{p \times p}^{\ge}$, we have $z \in \mathcal{N}(\widehat{
  \Sigma})$ and $z \in \mathcal{N}(\widehat{\Sigma}_k)$. In particular,
we have that $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset
\mathcal{N}(\widehat{\Sigma})$. Now, suppose $z \in
\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z^{T}
\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z^{T} \widehat{
  \Sigma}_k z$, which implies that $z \in \mathcal{N}(\widehat{
  \Sigma}_k)$ because $\widehat{\Sigma}_k \in \mathbb{R}_{p \times
  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\Sigma}) \subset
\mathcal{N}(\widehat{\Sigma}_k)$.
\end{proof}

In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
expression for $\tilde{\Sigma}_k$ in terms of the matrix of eigenvectors of
$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U^{T}$ be the
eigendecomposition of $\widehat{\Sigma}$ such that $D \in \mathbb{R}_{p
  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\Sigma}$
with $D = D_q \oplus 0_{p-q}$, $D_q \in \mathbb{R}_{q \times
  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
$\widehat{\Sigma}$, the columns of $U \in \mathbb{R}_{p \times p}$ are
the corresponding orthonormal eigenvectors of $\widehat{\Sigma}$, and
rank$(\widehat{\Sigma}) = q$. Then, we partition $U = (U_1, 
U_2)$ such that $U_1 \in \mathbb{R}_{p \times q}$ and $U_2 \in
\mathbb{R}_{p \times (p - q)}$.

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\Sigma} = U D U^{T}$ be the eigendecomposition of
$\widehat{\Sigma}$ as above, and suppose that rank$(\widehat{\Sigma}) =
q \le p$. Then, we have
\begin{align}
	\tilde{\Sigma}_k &= U(W_k \oplus \gamma I_{p-q})U^{T} \quad
  (k = 1, \ldots, K),\label{eq:rda-matrix}
\intertext{where}
W_k &= \alpha_k \{(1 - \lambda) U_1^{T} \widehat{\Sigma}_k U_1 + \lambda D_q\} + \gamma I_{q}.\label{eq:Wk}
\end{align}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $U_2$ span the null space
of $\widehat{\Sigma}$, which implies that $\widehat{\Sigma}_k U_2 =
0_{p \times (p - q)}$. Hence, $U^{T} \widehat{\Sigma}_k U = 
U_1^{T} \widehat{\Sigma}_k U_1 \oplus 0_{p-q}$ $(k = 1, \ldots,
K)$. Thus, $U^{T} \tilde{\Sigma}_k U = \alpha_k \{(1 - \lambda) U^{T}
\widehat{\Sigma}_k U + \lambda D\} + \gamma I_p$, and
\eqref{eq:rda-matrix} follows because $U$ is orthogonal.
\end{proof}

As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we
have the following corollary.

\begin{cor}
Let $\widehat{\Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
  rank$\{\widehat{\Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
\end{cor}
\begin{proof}
The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

Thus, by incorporating each $x_i$ into the estimation of $\Sigma_k$, we
increase the rank of $\widehat{\Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
0$. Next, we provide an essential result that enables us to prove that
\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
  \Sigma}_k$ corresponding to the null space of $\widehat{\Sigma}$ for any
unlabeled observation $x \in \mathbb{R}_{p \times 1}$.

\begin{lemma}\label{lemma:RDA-constant-term}
Let $U_2$ be defined as above. Then, for all $x \in \mathbb{R}_{p \times
  1}$, $U_2^{T} (x - \xbar_k) = U_2^{T} (x - \xbar_{k'})$ $(1 \le k,
k' \le K)$.
\end{lemma}
\begin{proof}
Let $x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
Recall that $U_2 \in \mathcal{N}(\widehat{\Sigma})$, which implies that
$U_2^{T} \in \mathcal{C}(\widehat{\Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $x_i \in \mathcal{C}(\widehat{
  \Sigma})$ $(i = 1, \ldots, N)$, $U_2^{T} x_i = 0_{(p-q) \times 1}$.
Hence, $0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i U_2^{T}x_i = U_2^{T}(\xbar_k -
\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
k') n_k \}$. Therefore, $U_2^{T} (x - \xbar_k) = U_2^{T} (x - \xbar_{k'})$.
\end{proof}

We now present our main result, where we decompose the decision rule of our
proposed classifier and show that the term requiring the largest computational
costs does not contribute to the classification of an unlabeled observation
performed using Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce
\eqref{eq:hdrda} to an equivalent, more computationally efficient decision rule.

>>>>

\title{Supplementary Material for ``Another Look at Regularized Discriminant Analysis for High-Dimensional Classification''}

@@

opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE,
               warning = FALSE, cache = TRUE, echo = FALSE, autodep = TRUE,
               fig.width = 12, fig.height = 12, out.width = "\\linewidth")
@

@@

@@

\begin{figure}
<<optimal_estimates_burczynski>>=
hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
p <- ggplot(subset(hdrda_optimal, data_set == "burczynski" & d %in% c(100, 500, 1000)), aes(lambda))
p <- p + geom_bar(aes(fill = gamma, weight = value))
p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
p <- p + theme(strip.text.y = element_text(size = 16))
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 24))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + scale_fill_discrete(name = expression(gamma))
p <- p + theme(legend.title = element_text(size = 28))
p <- p + theme(legend.key.size = unit(1.25, "cm"))
p <- p + theme(legend.text = element_text(size = 18))
p + xlab(expression(lambda)) + ylab("Counts")
@ 
\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
  classifier for the Burczynski data set for $d = 100$, $500$, and $1000$.}
\label{fig:optimal-burczynski}
\end{figure}

\begin{figure}
<<optimal_estimates_nakayama>>=
hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
p <- ggplot(subset(hdrda_optimal, data_set == "nakayama" & d %in% c(100, 500, 1000)), aes(lambda))
p <- p + geom_bar(aes(fill = gamma, weight = value))
p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
p <- p + theme(strip.text.y = element_text(size = 16))
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 24))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + scale_fill_discrete(name = expression(gamma))
p <- p + theme(legend.title = element_text(size = 28))
p <- p + theme(legend.key.size = unit(1.25, "cm"))
p <- p + theme(legend.text = element_text(size = 18))
p + xlab(expression(lambda)) + ylab("Counts")
@ 
\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
  classifier for the Nakayama data set for $d = 100$, $500$, and $1000$.}
\label{fig:optimal-nakayama}
\end{figure}

\begin{figure}
<<optimal_estimates_shipp>>=
hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
p <- ggplot(subset(hdrda_optimal, data_set == "shipp" & d %in% c(100, 500, 1000)), aes(lambda))
p <- p + geom_bar(aes(fill = gamma, weight = value))
p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
p <- p + theme(strip.text.y = element_text(size = 16))
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 24))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + scale_fill_discrete(name = expression(gamma))
p <- p + theme(legend.title = element_text(size = 28))
p <- p + theme(legend.key.size = unit(1.25, "cm"))
p <- p + theme(legend.text = element_text(size = 18))
p + xlab(expression(lambda)) + ylab("Counts")
@ 
\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
  classifier for the Shipp data set for $d = 100$, $500$, and $1000$.}
\label{fig:optimal-shipp}
\end{figure}

\begin{figure}
<<optimal_estimates_singh>>=
hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
p <- ggplot(subset(hdrda_optimal, data_set == "singh" & d %in% c(100, 500, 1000)), aes(lambda))
p <- p + geom_bar(aes(fill = gamma, weight = value))
p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
p <- p + theme(strip.text.y = element_text(size = 16))
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 24))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + scale_fill_discrete(name = expression(gamma))
p <- p + theme(legend.title = element_text(size = 28))
p <- p + theme(legend.key.size = unit(1.25, "cm"))
p <- p + theme(legend.text = element_text(size = 18))
p + xlab(expression(lambda)) + ylab("Counts")
@ 
\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
  classifier for the Singh data set for $d = 100$, $500$, and $1000$.}
\label{fig:optimal-singh}
\end{figure}
